\documentclass[12pt]{article}
\title{\rightline {\Huge {Due: Wed. Sept 20}}}
\author{\LARGE {HMC\quad Math 142 \quad Fall 2017} 
\\ {Prof. Gu}  
\\ {\LARGE Problem Set 3}}
\date{Start this assignment before Sunday night!}


%height and width adjustments

\setlength{\topmargin}{-1cm}
\setlength{\textwidth}{18cm}
\setlength{\textheight}{23cm}
\setlength{\oddsidemargin}{-1cm}

\begin{document}



\maketitle

\section*{ Read: } 

\begin{itemize}
\item{Baby Do Carmo, Differential Geometry
    of Curves and Surfaces:  
Sections 1-6, 1-7, Chapter 1}
\item{Handout 3}
\item{Lecture Notes}
\end{itemize}

\section*{ Do: }
\paragraph{A: Problems on Reviewing of Orthogonal transformations, 
Rotations, Reflections and Rigid Motions
  in $R^n$.}
\begin{itemize}
 
{\item a) Problem 6 on page 23, Section 1-5, Baby Do Carmo.}
\end{itemize}

\medskip
\paragraph{B: Problems from Lectures}

\begin{itemize}
{\item a) (Extra Credit: If you are familiar with groups) Show SO(n) is a group with respect to the usual matrix 
multiplication. (Later, we will see that SO(n) is in fact a Lie group.)}
{\item b) (Extra Credit:) Show that the mirror reflection $\tau$ (as defined in the 
lecture) is an orthogonal transformation and $\tau^2 = id$, where 
$id$ is the identity transformation. }
\end{itemize}


\paragraph{C: Other Problems}

\

\noindent {\bf Finish the two problems covered in class}
\begin{itemize}
	{\item a) Show that the shortest path between two points $p$ and $q$ is the straight line segment.}
	{\item b) Assume that all normals of a parametrized curve pass through a fixed point. Prove that the trace of the curve is contained in a circle. }
\end{itemize}


\noindent {\bf Choose 3 problems out of following problems:}
\begin{itemize}
{\item a) Problem 1 on page 22, Section 1-5, Baby Do Carmo.}
{\item b) Problem 2 on page 22, Section 1-5, Baby Do Carmo.}
{\item c) Problem 5 on page 23, Section 1-5, Baby Do Carmo.}
{\item d) Problem 12 on page 25, Section 1-5, Baby Do Carmo.}
\end{itemize}


\end{document}







