\documentclass[12pt]{article}
\title{\rightline {\Huge {Due: Wednesday, October 11}}}
\author{\LARGE {HMC\quad Math 142 \quad Fall 2017} 
\\ {Prof. Gu}  
\\ {\LARGE Problem Set 6}}
\date{Start this assignment before Sunday night!}


%height and width adjustments

\setlength{\topmargin}{-1cm}
\setlength{\textwidth}{18cm}
\setlength{\textheight}{23cm}
\setlength{\oddsidemargin}{-1cm}

\begin{document}



\maketitle

\section*{ Read: } 

\begin{itemize}
\item{Baby Do Carmo, Differential Geometry
    of Curves and Surfaces:  
Sections 2-4, 2-5, 2-6 and Section 5-10 on Abstract surfaces (starting 
on page 425)} 
\item{Handouts 8 and 9}
\item{Lecture Notes}
\end{itemize}

\section*{ Do: Please choose one of the following}
\begin{enumerate}
    \item Choose a project of your preference. Write 2 pages of your work of the project that could be background sections, introduction, development or implementation of an algorithm or a model, some theorems, etc.  Identify at least one or a couple research articles.
    \item Choose to work on homework problems. Work on part B, and choose 3 problems that interest you from part C to write up. Keys for part C are attached in resources.
    \item Choose half and half. Choose two problems from choice 2 (you can either choose 2 problems from part C or you can choose 1 problem on part C and works out the problem in part B ) and one page of work on choice 1
\end{enumerate}



\medskip
\paragraph{B: Problems from Lectures}

\begin{itemize}
% {\item a) Let $S$ be a subset of $R^3$.  Show that $S$ is 
% regular surface if and only if S is locally diffeomorphic 
% to $R^2$.}

{\item a) Find five examples of regular surfaces such that each of 
them can be represented as a surface of revolution.  Write down
specifically for each example the generating curve, the rotation 
axis, and the parameterization (as a map) for the surface (including
the domain of the map).}

\end{itemize}

\paragraph{C: Problems to choose from}
\begin{itemize}
{\item a) Problem 10 on page 81, Section 2-3, Baby Do Carmo.}
{\item b) Problem 9 on page 89, Section 2-4, Baby Do Carmo.}
{\item c) Problem 15 on page 90, Section 2-4, Baby Do Carmo.}
{\item d) Problem 18 on page 90, Section 2-4, Baby Do Carmo.}
{\item e) Problem 1 on page 99, Section 2-5, Baby Do Carmo.}
{\item f) Problem 3 on page 99, Section 2-5, Baby Do Carmo.}
{\item g) Problem 9 on page 100, Section 2-5, Baby Do Carmo.}
\end{itemize}


% \paragraph{D: Extra Credit Problems}
% \begin{itemize}
% {\item a) Let $T\subset R^3$ be a torus of revolution with center in 
% $(0,0,0)\in R^3$ and let $A(x,y,z)=(-x,-y,-z)$.
% Let $K$ be the quotient space of the torus $T$ by the equivalence 
% relation $p\sim A(p)$.  Can you tell what surface $K$ is? }
% {\item b) Show that $K$ is a differentiable 2-dimensional manifold.}
% {\item c) Show that $K$ is non orientable in two different ways.} 
% \end{itemize}

\end{document}













